---
title: "EDA"
output:
  github_document:
    df_print: paged
    html_preview: false
    toc: true
    toc_depth: 2
---

# TODO:

EDA:

1.  ~~investigate correlations between vars~~
2.  ~~`TARGET`-vars correlation~~
3.  ~~duplicates among lines, meaning of country label and country prefix in var names~~
4.  ~~are NAs simultaneously present in several cols for the same row ?~~

# Import libraries

```{r}
library(readr)
library(dplyr)
library(tidyr)
library(purrr)
library(tibble)
library(Hmisc)
library(ggplot2)
library(corrplot)
library(PerformanceAnalytics)
```

# Load data:

```{r}
X_train_raw <- read.csv("../data/X_train.csv")
X_test_raw <- read.csv("../data/X_test.csv")
y_train <- read.csv("../data/y_train.csv")
```

Shape of the data sets:

```{r}
print("X train:", )
dim(X_train_raw)

print("X test:", )
dim(X_test_raw)

print("y train:", )
dim(y_train)
```

# EDA

## ID alignement in X and y

Are samples' IDs in X and y aligned? Number of misaligned rows:

```{r}
sum(X_train_raw$ID != y_train$ID)
```

Yes, they are.

## Missing vals by variable

Missing values by column. Training set:

```{r}
tibble_ <- X_train_raw

na_cols_train <- tibble(
  column_names = names(tibble_),
  missing_percentage = round(colMeans(is.na(tibble_)) * 100, 1)
  ) %>% 
  arrange(desc(missing_percentage)) %>% 
  filter(missing_percentage > 0)

na_cols_train
```

Test set:

```{r}
tibble_ <- X_test_raw

na_cols_test <- tibble(
  column_names = names(tibble_),
  missing_percentage = round(colMeans(is.na(tibble_)) * 100, 1)
  ) %>% 
  arrange(desc(missing_percentage)) %>% 
  filter(missing_percentage > 0)

na_cols_test
```

Are cols with NAs the same in training and test sets?

```{r}
print("In train set only:")
print(setdiff(na_cols_train$column_names, na_cols_test$column_names))

print("In test set only:")
print(setdiff(na_cols_test$column_names, na_cols_train$column_names))
```

## Variables dtypes

Columns' dtypes. The dtypes of all columns have been properly detected on the data import.

```{r}
str(X_train_raw)
```

## `COUNTRY` var and country label prefix

Value counts in the `COUNTRY` column:

```{r}
X_train_raw %>%
  count(COUNTRY) %>%
  mutate(percentage = round(n / sum(n) * 100, 1))
```

The data describes two countries, FR and DE, roughly equally presented.

Duplicates among `DAY_ID` in the training set:

```{r}
X_train_raw %>%
  count(DAY_ID) %>%
  mutate(n = n - 1) %>% 
  rename(num_of_duplicates = n) %>% 
  count(num_of_duplicates) %>% 
  rename(count = n) %>% 
  mutate(percentage = round(count / sum(count) * 100, 1))
```

Most of the lines (3/4) have duplicates.

643\*2 + 208 gives the number of rows in the training set.

Duplicates among `DAY_ID` in the test set:

```{r}
X_test_raw %>%
  count(DAY_ID) %>%
  mutate(n = n - 1) %>% 
  rename(num_of_duplicates = n) %>% 
  count(num_of_duplicates) %>% 
  rename(count = n) %>% 
  mutate(percentage = round(count / sum(count) * 100, 1))
```

The number of duplicates is more or less the same as in the training set.

For the rows that have the same `DAY_ID`, in which columns are they different? Training set:

```{r}
tibble_ <- X_train_raw
tibble_$TARGET <- y_train$TARGET

# given a DAY_ID, select rows corresponding to it and get the names of the
# columns where the two rows having this DAY_ID differ
non_duplicated_cols <- function(day_id_) {
  col_names <- tibble_ %>% 
    filter(DAY_ID == day_id_) %>% 
    select_if(function(column) !identical(column[1], column[2])) %>% 
    names() %>% 
    paste(., collapse = ", ")
  return(col_names)
}

# get the list of day IDs that are duplicated in the table
ids <- tibble_ %>%
  count(DAY_ID) %>% 
  filter(n > 1) %>% 
  pull(DAY_ID)

# for each pair of duplicated rows, get a list of column names that have 
# different values in the pair of duplicated rows, drop duplicates
results <- lapply(ids, non_duplicated_cols) %>% unique()

results
```

Same question for the test set (no `TARGET`):

```{r}
tibble_ <- X_test_raw

# given a DAY_ID, select rows corresponding to it and get the names of the
# columns where the two rows having this DAY_ID differ
non_duplicated_cols <- function(day_id_) {
  col_names <- tibble_ %>% 
    filter(DAY_ID == day_id_) %>% 
    select_if(function(column) !identical(column[1], column[2])) %>% 
    names() %>% 
    paste(., collapse = ", ")
  return(col_names)
}

# get the list of day IDs that are duplicated in the table
ids <- tibble_ %>%
  count(DAY_ID) %>% 
  filter(n > 1) %>% 
  pull(DAY_ID)

# for each pair of duplicated rows, get a list of column names that have 
# different values in the pair of duplicated rows, drop duplicates
results <- lapply(ids, non_duplicated_cols) %>% unique()

results
```

Thus the rows having the same value of `DAY_ID` differ only by `ID` (obviously), `COUNTRY` (FR or DE) and `TARGET` (prediction for FR or DE). For the rest they are identical.

## Analysis of the lines with and without `DAY_ID` duplicates.

First select rows that don't have duplicates:

```{r}
# get the list of day IDs that are not duplicated in the training set
ids_train <- X_train_raw %>%
  count(DAY_ID) %>% 
  filter(n == 1) %>% 
  pull(DAY_ID)

# get the list of day IDs that are not duplicated in the test set
ids_test <- X_test_raw %>%
  count(DAY_ID) %>% 
  filter(n == 1) %>% 
  pull(DAY_ID)
```

To which countries do these lines in the training and test sets correspond?

```{r}
print(X_train_raw %>% filter(DAY_ID %in% ids_train) %>% count(COUNTRY))
print(X_test_raw %>% filter(DAY_ID %in% ids_test) %>% count(COUNTRY))
```

All rows that don't have a duplicate in `DAY_ID` correspond to country FR.

What about missing vals in these non duplicated lines?

```{r}
tibble_ <- X_train_raw %>% filter(DAY_ID %in% ids_train)

na_cols_train <- tibble(
  column_names = names(tibble_),
  missing_percentage = round(colMeans(is.na(tibble_)) * 100, 1)
  ) %>% 
  arrange(desc(missing_percentage)) %>% 
  filter(missing_percentage > 0)

print(na_cols_train)

tibble_ <- X_test_raw %>% filter(DAY_ID %in% ids_test)

na_cols_test <- tibble(
  column_names = names(tibble_),
  missing_percentage = round(colMeans(is.na(tibble_)) * 100, 1)
  ) %>% 
  arrange(desc(missing_percentage)) %>% 
  filter(missing_percentage > 0)

print(na_cols_test)
```

To conclude, the lines that don't have duplicates in `DAY_ID` have NAs in the cols describing countries' net import and export and DE\<-\>FR exchange (`DE_FR_EXCHANGE`, `FR_DE_EXCHANGE`, `DE_NET_EXPORT`, `FR_NET_EXPORT`, `DE_NET_IMPORT`, `FR_NET_IMPORT`).

And in duplicated lines?

```{r}
tibble_ <- X_train_raw %>% filter(!(DAY_ID %in% ids_train))

na_cols_train <- tibble(
  column_names = names(tibble_),
  missing_percentage = round(colMeans(is.na(tibble_)) * 100, 1)
  ) %>% 
  arrange(desc(missing_percentage)) %>% 
  filter(missing_percentage > 0)

print(na_cols_train)

tibble_ <- X_test_raw %>% filter(!(DAY_ID %in% ids_test))

na_cols_test <- tibble(
  column_names = names(tibble_),
  missing_percentage = round(colMeans(is.na(tibble_)) * 100, 1)
  ) %>% 
  arrange(desc(missing_percentage)) %>% 
  filter(missing_percentage > 0)

print(na_cols_test)
```

The lines having duplicates in `DAY_ID` have NAs only in the cols describing countries' weather conditions (`DE_RAIN`, `FR_RAIN`, `DE_WIND`, `FR_WIND`, `DE_TEMP`, `FR_TEMP`).

The two lists of columns above span all (12) the columns with missing values.

## Are NAs simultaneously present in several cols for the same row ?

Find the groups of columns that have simultaneously NAs in one row for the training set:

```{r}
missing_columns_list <- X_train_raw %>%
  pmap(function(...) names(list(...))[sapply(list(...), is.na)]) %>% 
  Filter(function(x) length(x) > 0, .) %>% 
  unique(.)

missing_columns_list
```

Same for the test set:

```{r}
X_test_raw %>%
  pmap(function(...) names(list(...))[sapply(list(...), is.na)]) %>% 
  Filter(function(x) length(x) > 0, .) %>% 
  unique(.)
```

Apparently we get the same groups of columns.

As a side note, the analysis above shows that the test set has the same properties as the training set and thus it is a representative sample of the data.

To conclude the analysis above:

-   The rows that **don't have** duplicates in the `DAY_ID` col have NAs in all 6 cols describing countries' weather conditions (`DE_RAIN`, `FR_RAIN`, `DE_WIND`, `FR_WIND`, `DE_TEMP`, `FR_TEMP`).
-   The rows that **do have** duplicates in the `DAY_ID` col can have NAs simultaneously in the following groups of cols:
    -   `DE_NET_EXPORT`, `DE_NET_IMPORT`
    -   `DE_NET_EXPORT`, `FR_NET_EXPORT`, `DE_NET_IMPORT`, `FR_NET_IMPORT`
    -   `DE_FR_EXCHANGE`, `FR_DE_EXCHANGE`, `DE_NET_EXPORT`, `FR_NET_EXPORT`, `DE_NET_IMPORT`, `FR_NET_IMPORT`

## Correlation between vars

Excluding vars `ID` and `DAY_ID` useless for the prediction and setting apart `COUNTRY`, we are left with numeric vars only.

Pair corr matrix. I use Spearman corr to capture non-linear corrs.

```{r}
res <- cor(select(X_train_raw, -ID, -DAY_ID, -COUNTRY), method = "spearman", use = "complete.obs")

res2 <- rcorr(as.matrix(select(X_train_raw, -ID, -DAY_ID, -COUNTRY)), type = "spearman")
```

Plot:

```{r}
corrplot(
  res,
  type = "upper",
  order = "hclust", 
  tl.col = "black",
  tl.srt = 45,
  tl.cex = 0.4)
```

Let's explore in more details the correlation matrix. For the convenience we will flatten it together with the matrix of p-values.

```{r}
# Code source: http://www.sthda.com/english/wiki/correlation-matrix-a-quick-start-guide-to-analyze-format-and-visualize-a-correlation-matrix-using-r-software#a-simple-function-to-format-the-correlation-matri
# ++++++++++++++++++++++++++++
# flattenCorrMatrix
# ++++++++++++++++++++++++++++
# cormat : matrix of the correlation coefficients
# pmat : matrix of the correlation p-values
flattenCorrMatrix <- function(cormat, pmat) {
  ut <- upper.tri(cormat)
  data.frame(
    row = rownames(cormat)[row(cormat)[ut]],
    column = rownames(cormat)[col(cormat)[ut]],
    cor  =(cormat)[ut],
    p = pmat[ut]
    )
}

res2_flat <- flattenCorrMatrix(res2$r, res2$P) %>% 
  mutate(cor = round(cor, 2), p = round(100 * p, 0)) %>% 
  arrange(desc(abs(cor)))

res2_flat
```

First we notice that all correlations with an absolute value greater than 0.07 are significant with p-value \< 1%.

Other observations:

1.  The following pairs of variables are all perfectly anti correlated: `DE_FR_EXCHANGE` vs `FR_DE_EXCHANGE`, `DE_NET_EXPORT` vs `DE_NET_IMPORT`, `FR_NET_EXPORT` vs `FR_NET_IMPORT`. Actually one variable in the pair is the inverse of another. The plot below shows all pair correlations between these variables. From it , and from the correlation matrix, we also notice that DE\<-\>FR exchange is strongly correlated with DE/FR net import/export. At the same time `DE_NET_IMPORT` and `FR_NET_IMPORT` are decorrelated.

```{r}
chart.Correlation(
  select(X_train_raw, DE_FR_EXCHANGE, FR_DE_EXCHANGE, DE_NET_EXPORT, DE_NET_IMPORT, FR_NET_EXPORT, FR_NET_IMPORT), 
  histogram=TRUE, 
  pch=19)

chart.Correlation(
  select(X_train_raw, FR_DE_EXCHANGE, DE_NET_IMPORT, FR_NET_EXPORT), 
  histogram=TRUE, 
  pch=19)
```

Relation between FR/DE export, FR\<-\>DE exchange and FR/DE consumption (total and without renewable).

-   Notice the difference in the relation between `FR_CONSUMPTION` and `FR_RESIDUAL_LOAD` versus `DE_CONSUMPTION` and `DE_RESIDUAL_LOAD`. Could it be attributed to the different proportion of the renewable sources in two countries?
-   Weak but noticeable relation between e.g. `FR_DE_EXCHANGE` and `FR_CONSUMPTION`.

```{r}
chart.Correlation(
  select(X_train_raw, FR_DE_EXCHANGE, DE_NET_EXPORT, FR_NET_EXPORT, FR_CONSUMPTION, DE_CONSUMPTION, FR_RESIDUAL_LOAD, DE_RESIDUAL_LOAD), 
  histogram=TRUE, 
  pch=19)
```

Relation between weather conditions in two countries. Rather strong correlation between temperature and wind in FR and DE, the amount of rain in two countries is essentially unrelated.

```{r}
chart.Correlation(
  select(X_train_raw, DE_RAIN, FR_RAIN, DE_WIND, FR_WIND, DE_TEMP, FR_TEMP), 
  histogram=TRUE, 
  pch=19)
```

Relation between weather conditions and production of renewable energy in two countries.

-   An odd relation between `x_WIND` and `x_WINDPOW` for both countries, as it consisted of two approx linear relations with different slopes.
-   For both countries, there may be a steep anti correlation between amount of rain and solar energy production. However, the distribution of x_RAIN values has a long high-end wing that may hinder this relation.

```{r}
chart.Correlation(
  select(X_train_raw, FR_RAIN, FR_WIND, FR_TEMP, FR_SOLAR, FR_WINDPOW, FR_HYDRO), 
  histogram=TRUE, 
  pch=19)

chart.Correlation(
  select(X_train_raw, DE_RAIN, DE_WIND, DE_TEMP, DE_SOLAR, DE_WINDPOW, DE_HYDRO), 
  histogram=TRUE, 
  pch=19)
```

Relations between countries' consumption rates, exchange between them and whether conditions.

No clear relationship between any of the pair of variables.

```{r}
chart.Correlation(
  select(X_train_raw, FR_CONSUMPTION, FR_RESIDUAL_LOAD, FR_DE_EXCHANGE, FR_NET_EXPORT, FR_RAIN, FR_WIND, FR_TEMP), 
  histogram=TRUE, 
  pch=19)
```

Correlation between consumption and production.

```{r}
chart.Correlation(
  select(X_train_raw, FR_CONSUMPTION, FR_RESIDUAL_LOAD, FR_GAS, FR_COAL, FR_NUCLEAR), 
  histogram=TRUE, 
  pch=19)
```

## Relations between explanatory vars and target

We exclude from the analysis `ID`, `DAY_ID` and perfectly correlated vars `DE_FR_EXCHANGE`, `FR_NET_IMPORT`, `DE_NET_IMPORT`.

For both countries there is no clear relationship between any of the explanatory variable and target.

For France:

```{r}
# Merge the two tibbles based on a common identifier
merged_tibble <- left_join(X_train_raw, y_train, by = "ID") %>% 
  filter(COUNTRY == "FR") %>% 
  select(-ID, -DAY_ID, -COUNTRY, -DE_FR_EXCHANGE, -FR_NET_IMPORT, -DE_NET_IMPORT)

# Create a vector of explanatory variable names
explanatory_vars <- colnames(select(merged_tibble, -TARGET))

# Generate scatter plots for each explanatory variable
for (var in colnames(select(merged_tibble, -TARGET))) {
  res <- ggplot(merged_tibble, aes_string(x = var, y = "TARGET")) +
    geom_point() +
    geom_smooth(method = "loess", se = FALSE, color = "red", span = 0.9) +
    labs(x = var, y = "Target Variable") +
    ggtitle(paste("Scatter Plot of", var, "vs. Target Variable for FR"))
  print(res)
}

```

For Germany:

```{r}
# Merge the two tibbles based on a common identifier
merged_tibble <- left_join(X_train_raw, y_train, by = "ID") %>% 
  filter(COUNTRY == "DE") %>% 
  select(-ID, -DAY_ID, -COUNTRY, -DE_FR_EXCHANGE, -FR_NET_IMPORT, -DE_NET_IMPORT)

# Create a vector of explanatory variable names
explanatory_vars <- colnames(select(merged_tibble, -TARGET))

# Generate scatter plots for each explanatory variable
for (var in colnames(select(merged_tibble, -TARGET))) {
  res <- ggplot(merged_tibble, aes_string(x = var, y = "TARGET")) +
    geom_point() +
    geom_smooth(method = "loess", se = FALSE, color = "red", span = 0.9) +
    labs(x = var, y = "Target Variable") +
    ggtitle(paste("Scatter Plot of", var, "vs. Target Variable for DE"))
  print(res)
}

```

Both countries together:

```{r}
# Merge the two tibbles based on a common identifier
merged_tibble <- left_join(X_train_raw, y_train, by = "ID") %>% 
  select(-ID, -DAY_ID, -COUNTRY, -DE_FR_EXCHANGE, -FR_NET_IMPORT, -DE_NET_IMPORT)

# Create a vector of explanatory variable names
explanatory_vars <- colnames(select(merged_tibble, -TARGET))

# Generate scatter plots for each explanatory variable
for (var in colnames(select(merged_tibble, -TARGET))) {
  res <- ggplot(merged_tibble, aes_string(x = var, y = "TARGET")) +
    geom_point() +
    geom_smooth(method = "loess", se = FALSE, color = "red", span = 0.9) +
    labs(x = var, y = "Target Variable") +
    ggtitle(paste("Scatter Plot of", var, "vs. Target Variable for DE"))
  print(res)
}
```

## Value distribution for the explanatory and target vars

For France.

Notice the distribution of `x_CONSUMPTION`, `x_COAL`, `x_GAZ` and `x_NUCLEAR` and the difference in these distribs for the two countries.

```{r}
# Merge the two tibbles based on a common identifier
merged_tibble <- X_train_raw %>% 
  filter(COUNTRY == "FR") %>% 
  select(-ID, -DAY_ID, -COUNTRY, -DE_FR_EXCHANGE, -FR_NET_IMPORT, -DE_NET_IMPORT)

# Generate hist for each explanatory variable
for (var in colnames(merged_tibble)) {
  res <- ggplot(merged_tibble, aes_string(x = var)) +
    geom_histogram() +
    ggtitle(paste(var, " for FR"))
  print(res)
}
```

For Germany only the TARGET distribution is different, all other cols are identical to France:

```{r}
# Merge the two tibbles based on a common identifier
merged_tibble <- left_join(X_train_raw, y_train, by = "ID") %>% 
  filter(COUNTRY == "DE") %>% 
  select(-ID, -DAY_ID, -COUNTRY, -DE_FR_EXCHANGE, -FR_NET_IMPORT, -DE_NET_IMPORT)

res <- ggplot() +
  geom_histogram(aes(x = y_train[X_train_raw$COUNTRY == "FR", "TARGET"], y = stat(count / sum(count))), fill = "blue", alpha = 0.5) +
  geom_histogram(aes(x = y_train[X_train_raw$COUNTRY == "DE", "TARGET"], y = stat(count / sum(count))), fill = "red", alpha = 0.5) +
  labs(title = "TARGET for FR and DE", x = "TARGET") + 
  scale_fill_manual(values = c("blue", "red"), labels = c("TARGET FR", "TARGET DE"))
print(res)

```

## TARGET FR vs TARGET DE

No clear relation between target values for the same day for the two countries.

```{r}
rearranged_data <- left_join(X_train_raw, y_train, by = "ID") %>% 
  select(DAY_ID, COUNTRY, TARGET) %>%
  pivot_wider(names_from = COUNTRY, values_from = TARGET, values_fill = NA) %>%
  drop_na()

# Display the rearranged table
rearranged_data

res <- ggplot(rearranged_data) +
  geom_point(aes(x = FR, y = DE)) +
  labs(x = "TARGET FR", y = "TARGET DE")
print(res)
```

## TARGET distribution for missing and non-missing vals

We consider all 12 variables that have missing values and plot a histogram of TARGET values for each of them for the rows with and without missing values. We plot these histograms separately for FR and DE.

First, for FR. Apparently the presence of missing values is not related to the TARGET value.

```{r}
df <- X_train_raw %>% 
  filter(COUNTRY == "FR") %>% 
  select(
    FR_DE_EXCHANGE, 
    FR_NET_EXPORT, 
    DE_NET_IMPORT, 
    DE_RAIN, 
    FR_RAIN, 
    DE_WIND, 
    FR_WIND, 
    DE_TEMP, 
    FR_TEMP)

# Generate hist for each explanatory variable
for (var in colnames(df)) {
  mask <- df %>% select(., all_of(var)) %>% is.na()
  res <- ggplot() +
    geom_histogram(aes(x = y_train[!mask, "TARGET"], y = stat(count / sum(count))), fill="blue", alpha = 0.5) +
    geom_histogram(aes(x = y_train[mask, "TARGET"], y = stat(count / sum(count))), fill="red", alpha = 0.5) +
    ggtitle(paste(var, " for FR"))
  print(res)
}
```

For DE. SAme conclusion: no clear relation between the presence of missing values and TARGET. Note that the absence of missing values for `FR_DE_EXCHANGE`, `FR_NET_EXPORT`, `DE_NET_IMPORT` is in line with the previous analysis (DE lines always have a duplicate `DAY_ID` for FR).

```{r}
df <- X_train_raw %>% 
  filter(COUNTRY == "DE") %>% 
  select(
    FR_DE_EXCHANGE, 
    FR_NET_EXPORT, 
    DE_NET_IMPORT, 
    DE_RAIN, 
    FR_RAIN, 
    DE_WIND, 
    FR_WIND, 
    DE_TEMP, 
    FR_TEMP)

# Generate hist for each explanatory variable
for (var in colnames(df)) {
  mask <- df %>% select(., all_of(var)) %>% is.na()
  res <- ggplot() +
    geom_histogram(aes(x = y_train[!mask, "TARGET"], y = stat(count / sum(count))), fill="blue", alpha = 0.5) +
    geom_histogram(aes(x = y_train[mask, "TARGET"], y = stat(count / sum(count))), fill="red", alpha = 0.5) +
    ggtitle(paste(var, " for DE"))
  print(res)
}
```

## Feature interactions

Heat maps with color code corresponding to the mean TARGET per bin and text - number of examples per bin.

```{r}
plot_var_interaction <- function(df, col1, col2, country, target = "TARGET", nbins = 5) {
  
  # Create bins and calculate mean TARGET and count for each bin
  summary_data <- df %>%
    filter(COUNTRY == country) %>% 
    select(col1, col2, target) %>% 
    mutate(
      col1_bin = cut(!!sym(col1), breaks = nbins),
      col2_bin = cut(!!sym(col2), breaks = nbins)
    ) %>%
    group_by(col1_bin, col2_bin) %>%
    summarise(mean_target = mean(!!sym(target), na.rm = TRUE), count_target = n())
  
  # Draw the two-dimensional histogram with colored bins
  result <- ggplot(summary_data, aes(x = col1_bin, y = col2_bin)) +
    geom_tile(aes(fill = mean_target)) +
    geom_text(aes(label = count_target), color = "white", size = 4) +
    scale_fill_gradient(low = "blue", high = "red")  +
    labs(x = col1, y = col2)

  print(result)
}
```

There is no visible trend in the interaction between `FR_NUCLEAR` and `FR_CONSUMPTION`:

```{r}
col1 <- "FR_NUCLEAR"
col2 <- "FR_CONSUMPTION"
country = "FR"
plot_var_interaction(left_join(X_train_raw, y_train, by = "ID"), col1, col2, country)

col1 <- "DE_NUCLEAR"
col2 <- "DE_CONSUMPTION"
country = "DE"
plot_var_interaction(left_join(X_train_raw, y_train, by = "ID"), col1, col2, country)
```

Nor there is any trend in the interaction between `FR_NUCLEAR` and `FR_RESIDUAL_LOAD`:

```{r}
col1 <- "FR_NUCLEAR"
col2 <- "FR_RESIDUAL_LOAD"
country = "FR"
plot_var_interaction(left_join(X_train_raw, y_train, by = "ID"), col1, col2, country)

col1 <- "DE_NUCLEAR"
col2 <- "DE_RESIDUAL_LOAD"
country = "DE"
plot_var_interaction(left_join(X_train_raw, y_train, by = "ID"), col1, col2, country)
```

```{r}
col1 <- "FR_GAS"
col2 <- "FR_RESIDUAL_LOAD"
country = "FR"
plot_var_interaction(left_join(X_train_raw, y_train, by = "ID"), col1, col2, country)

col1 <- "DE_GAS"
col2 <- "DE_RESIDUAL_LOAD"
country = "DE"
plot_var_interaction(left_join(X_train_raw, y_train, by = "ID"), col1, col2, country)
```

```{r}
col1 <- "FR_COAL"
col2 <- "FR_RESIDUAL_LOAD"
country = "FR"
plot_var_interaction(left_join(X_train_raw, y_train, by = "ID"), col1, col2, country)

col1 <- "DE_COAL"
col2 <- "DE_RESIDUAL_LOAD"
country = "DE"
plot_var_interaction(left_join(X_train_raw, y_train, by = "ID"), col1, col2, country)
```

```{r}
col1 <- "FR_DE_EXCHANGE"
col2 <- "FR_RESIDUAL_LOAD"
country = "FR"
plot_var_interaction(left_join(X_train_raw, y_train, by = "ID"), col1, col2, country)

col1 <- "FR_DE_EXCHANGE"
col2 <- "DE_RESIDUAL_LOAD"
country = "DE"
plot_var_interaction(left_join(X_train_raw, y_train, by = "ID"), col1, col2, country)
```

```{r}
col1 <- "FR_NUCLEAR"
col2 <- "FR_GAS"
country = "FR"
plot_var_interaction(left_join(X_train_raw, y_train, by = "ID"), col1, col2, country)

col1 <- "DE_NUCLEAR"
col2 <- "DE_GAS"
country = "DE"
plot_var_interaction(left_join(X_train_raw, y_train, by = "ID"), col1, col2, country)
```

```{r}
col1 <- "FR_NUCLEAR"
col2 <- "FR_COAL"
country = "FR"
plot_var_interaction(left_join(X_train_raw, y_train, by = "ID"), col1, col2, country)

col1 <- "DE_NUCLEAR"
col2 <- "DE_COAL"
country = "DE"
plot_var_interaction(left_join(X_train_raw, y_train, by = "ID"), col1, col2, country)
```

There is not enough stats to evidence the pattern visible in the interaction between `GAS_RET` and `CARBON_RET`:

```{r}
col1 <- "GAS_RET"
col2 <- "CARBON_RET"
country = "FR"
plot_var_interaction(left_join(X_train_raw, y_train, by = "ID"), col1, col2, country)

country = "DE"
plot_var_interaction(left_join(X_train_raw, y_train, by = "ID"), col1, col2, country)
```

Same for the interaction between `COAL_RET` and `CARBON_RET`:

```{r}
col1 <- "COAL_RET"
col2 <- "CARBON_RET"
country = "FR"
plot_var_interaction(left_join(X_train_raw, y_train, by = "ID"), col1, col2, country)

country = "DE"
plot_var_interaction(left_join(X_train_raw, y_train, by = "ID"), col1, col2, country)
```

Idem for the interaction between `COAL_RET` and `GAS_RET`:

```{r}
col1 <- "COAL_RET"
col2 <- "GAS_RET"
country = "FR"
plot_var_interaction(left_join(X_train_raw, y_train, by = "ID"), col1, col2, country)

country = "DE"
plot_var_interaction(left_join(X_train_raw, y_train, by = "ID"), col1, col2, country)
```

Idem:

```{r}
col1 <- "FR_GAS"
col2 <- "GAS_RET"
country = "FR"
plot_var_interaction(left_join(X_train_raw, y_train, by = "ID"), col1, col2, country)

col1 <- "DE_GAS"
col2 <- "GAS_RET"
country = "DE"
plot_var_interaction(left_join(X_train_raw, y_train, by = "ID"), col1, col2, country)
```

Idem:

```{r}
col1 <- "FR_COAL"
col2 <- "COAL_RET"
country = "FR"
plot_var_interaction(left_join(X_train_raw, y_train, by = "ID"), col1, col2, country)

col1 <- "DE_COAL"
col2 <- "COAL_RET"
country = "DE"
plot_var_interaction(left_join(X_train_raw, y_train, by = "ID"), col1, col2, country)
```

```{r}
col1 <- "FR_GAS"
col2 <- "FR_NUCLEAR"
country = "FR"
target <- "TARGET"

# Merge the two tibbles based on a common identifier
merged_tibble <- left_join(X_train_raw, y_train, by = "ID") %>% 
  filter(COUNTRY == country) %>% 
  select(!!sym(col1), !!sym(col2), !!sym(target))

res <- ggplot(merged_tibble) +
  geom_point(aes(x = !!sym(col1), y = !!sym(col2), color = !!sym(target)), size = 1, alpha = 0.9) +
  scale_color_viridis_c("Taux de chômage", option = "plasma")
print(res)

col1 <- "FR_COAL"
col2 <- "FR_NUCLEAR"
country = "FR"
target <- "TARGET"

# Merge the two tibbles based on a common identifier
merged_tibble <- left_join(X_train_raw, y_train, by = "ID") %>% 
  filter(COUNTRY == country) %>% 
  select(!!sym(col1), !!sym(col2), !!sym(target))

res <- ggplot(merged_tibble) +
  geom_point(aes(x = !!sym(col1), y = !!sym(col2), color = !!sym(target)), size = 1, alpha = 0.9) +
  scale_color_viridis_c("Taux de chômage", option = "plasma")
print(res)

col1 <- "DE_GAS"
col2 <- "DE_NUCLEAR"
country = "DE"
target <- "TARGET"

# Merge the two tibbles based on a common identifier
merged_tibble <- left_join(X_train_raw, y_train, by = "ID") %>% 
  filter(COUNTRY == country) %>% 
  select(!!sym(col1), !!sym(col2), !!sym(target))

res <- ggplot(merged_tibble) +
  geom_point(aes(x = !!sym(col1), y = !!sym(col2), color = !!sym(target)), size = 1, alpha = 0.9) +
  scale_color_viridis_c("Taux de chômage", option = "plasma")
print(res)

col1 <- "DE_COAL"
col2 <- "DE_NUCLEAR"
country = "DE"
target <- "TARGET"

# Merge the two tibbles based on a common identifier
merged_tibble <- left_join(X_train_raw, y_train, by = "ID") %>% 
  filter(COUNTRY == country) %>% 
  select(!!sym(col1), !!sym(col2), !!sym(target))

res <- ggplot(merged_tibble) +
  geom_point(aes(x = !!sym(col1), y = !!sym(col2), color = !!sym(target)), size = 1, alpha = 0.9) +
  scale_color_viridis_c("Taux de chômage", option = "plasma")
print(res)
```

```{r}
col1 <- "FR_COAL"
col2 <- "COAL_RET"
country = "FR"
target <- "TARGET"

# Merge the two tibbles based on a common identifier
merged_tibble <- left_join(X_train_raw, y_train, by = "ID") %>% 
  filter(COUNTRY == country) %>% 
  select(!!sym(col1), !!sym(col2), !!sym(target))

res <- ggplot(merged_tibble) +
  geom_point(aes(x = !!sym(col1), y = !!sym(col2), color = !!sym(target)), size = 2, alpha = 0.9) +
  scale_color_viridis_c("Taux de chômage", option = "plasma")
print(res)

```

```{r}
col1 <- "FR_COAL"
col2 <- "CARBON_RET"
country = "FR"
target <- "TARGET"

# Merge the two tibbles based on a common identifier
merged_tibble <- left_join(X_train_raw, y_train, by = "ID") %>% 
  filter(COUNTRY == country) %>% 
  select(!!sym(col1), !!sym(col2), !!sym(target))

res <- ggplot(merged_tibble) +
  geom_point(aes(x = !!sym(col1), y = !!sym(col2), color = !!sym(target)), size = 2, alpha = 0.9) +
  scale_color_viridis_c("Taux de chômage", option = "plasma")
print(res)

```

## dTARGET as target

```{r}
day_ids <- X_train_raw%>% 
  filter(COUNTRY == "DE") %>% 
  pull(DAY_ID)

mask_fr_ids <- (X_train_raw$DAY_ID %in% day_ids) & (X_train_raw$COUNTRY == "FR")

mask_de_ids <- (X_train_raw$DAY_ID %in% day_ids) & (X_train_raw$COUNTRY == "DE")

merged_tibble <- X_train_raw[mask_fr_ids,] %>%
  select(-ID, -DAY_ID, -COUNTRY, -DE_FR_EXCHANGE, -FR_NET_IMPORT, -DE_NET_IMPORT)

merged_tibble$delta_TARGET <- y_train[mask_fr_ids, "TARGET"] - y_train[mask_de_ids, "TARGET"]

# Generate scatter plots for each explanatory variable
for (var in colnames(select(merged_tibble, -delta_TARGET))) {
  res <- ggplot(merged_tibble, aes_string(x = var, y = "delta_TARGET")) +
    geom_point() +
    geom_smooth(method = "loess", se = FALSE, color = "red", span = 0.9) +
    labs(x = var, y = "delta Target Variable") +
    ggtitle(paste("Scatter Plot of", var, "vs. delta Target Variable"))
  print(res)
}

```

# Reproduce benchmark

## Data preparation

Fill-in NAs:

```{r}
X_train <- X_train_raw %>% replace(is.na(.), 0)
X_test <- X_test_raw %>% replace(is.na(.), 0)
```

Drop unused cols. Mind that we use `ID` and `DAY_ID` as explanatory vars.

```{r}
X_train <- X_train %>% select(-COUNTRY)
X_test <- X_test %>% select(-COUNTRY)
```

## Model

Fit OLS:

```{r}
lm_ <- lm(y_train$TARGET ~ ., data = X_train)
```

## FEature importance

```{r}
coef(lm_) %>% 
  enframe(., name = "Names", value = "Values") %>% 
  arrange(desc(abs(Values)))
```

## Comparison of Py and R OLS

Compare with the OLS model in Py benchmlark notebook:

```{r}
py_lm_ <- read.csv("../tmp/py_benchmark_model_coeffs.csv")
r_lm_ <- tibble(param = names(coef(lm_)), val = coef(lm_))

# rename Intercept
r_lm_$param[1] = "intercept"

# outer join between two tibbles
# ...
lm_join_ <- full_join(py_lm_, r_lm_, by = "param", suffix = c(".py", ".r"))

# element wise comparison of the models' coefficients
lm_join_$equal <- near(py_lm_$val, r_lm_$val)

lm_join_
```

Except few variables (`DE_FR_EXCHANGE` (different), `FR_DE_EXCHANGE` (NA), `DE_NET_EXPORT` (different),`FR_NET_EXPORT` (different), `DE_NET_IMPORT`(NA), `FR_NET_IMPORT` (NA)), the coefficients of the models are the same. For the coefficients that differ, the reason is that OLS in R has automatically detected highly correlated variables and excluded them from the model, thus some of the coefficients are NAs and others are different.

## Model prediction and export results

Predict:

```{r}
y_hat_train <- predict(lm_, X_train)
y_hat_test <- predict(lm_, X_test)
```

Spearman correlation for the predictions on the training set:

```{r}
cor(y_train$TARGET, y_hat_train, method = "spearman")
```

This correlation value is the same as the one of the Py benchmark model.

Export predictions on the test set:

```{r}
to_export <- tibble(ID = X_test_raw$ID, TARGET = y_hat_test)

write_csv(to_export, "../data/y_hat_test.csv")
```
