---
title: "EDA"
output:
  html_document:
    df_print: paged
---

```{r}
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
```

# Load data:

```{r}
X_train_raw <- read.csv("../data/X_train.csv")
X_test_raw <- read.csv("../data/X_test.csv")
y_train <- read.csv("../data/y_train.csv")
```

# EDA

Are samples' IDs in X and y aligned?

```{r}
sum(X_train_raw$ID != y_train$ID)
```

Missing values by column. Training set:

```{r}
tibble_ <- X_train_raw

na_cols_train <- tibble(
  column_names = names(tibble_),
  missing_percentage = round(colMeans(is.na(tibble_)) * 100, 1)
  ) %>% 
  arrange(desc(missing_percentage)) %>% 
  filter(missing_percentage > 0)

na_cols_train
```

Test set:

```{r}
tibble_ <- X_test_raw

na_cols_test <- tibble(
  column_names = names(tibble_),
  missing_percentage = round(colMeans(is.na(tibble_)) * 100, 1)
  ) %>% 
  arrange(desc(missing_percentage)) %>% 
  filter(missing_percentage > 0)

na_cols_test
```

Are cols with NAs the same in training and test sets?

```{r}
print("In train set only:")
print(setdiff(na_cols_train$name, na_cols_test$name))

print("In test set only:")
print(setdiff(na_cols_test$name, na_cols_train$name))
```

Columns' dtypes:

```{r}
str(X_train_raw)
```

# Reproduce benchmark

Fill-in NAs:

```{r}
X_train <- X_train_raw %>% replace(is.na(.), 0)
X_test <- X_test_raw %>% replace(is.na(.), 0)
```

Drop unused cols. Mind that we use `ID` and `DAY_ID` as explanatory vars.

```{r}
X_train <- X_train %>% select(-COUNTRY)
X_test <- X_test %>% select(-COUNTRY)
```

Fit OLS:

```{r}
lm_ <- lm(y_train$TARGET ~ ., data = X_train)
```

Compare with the OLS model in Py benchmlark notebook:

```{r}
py_lm_ <- read.csv("../tmp/py_benchmark_model_coeffs.csv")
r_lm_ <- tibble(param = names(coef(lm_)), val = coef(lm_))

# rename Intercept
r_lm_$param[1] = "intercept"

# outer join between two tibbles
# ...
lm_join_ <- full_join(py_lm_, r_lm_, by = "param", suffix = c(".py", ".r"))

# element wise comparison of the models' coefficients
lm_join_$equal <- near(py_lm_$val, r_lm_$val)

lm_join_
```

Predict:

```{r}
y_hat_train <- predict(lm_, X_train)
y_hat_test <- predict(lm_, X_test)
```

Spearman correlation for the predictions on the training set:

```{r}
cor(y_train$TARGET, y_hat_train, method = "spearman")
```

Export predictions on the test set:

```{r}
to_export <- tibble(ID = X_test_raw$ID, TARGET = y_hat_test)

write_csv(to_export, "../data/y_hat_test.csv")
```
